\textbf{Supplementary Information}

### 2020-09-16 submission/manuscript.Rmd
Pipeline optimization.

### Data processing and visualization
Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018]. The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).

>>>>

### Data processing and visualization
Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018]. The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).

### 2020-09-16 submission/manuscript.tex
Pipeline optimization.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

@@

\title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
Surfaces}}
\author{}
\date{\vspace{-2.5em}}

@@

To evaluate if the obtained DNA is suitable to determine the microbial
community structure an Illumina sequencing of the V4 16S rRNA region was
performed. Sequencing yielded a total of 336,937 sequences after quality
curation and exclusion of eukaryotic, mitochondrial and no relative
sequences. The number of sequences classified as chloroplast was 97,328.
After excluding these sequences the total number of retrieved reads
dropped to 239,609, ranging from 22,587 to 52,958 sequences per sample
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off which is commonly
observed in high-throughput 16S rRNA amplicon sequencing procedures
(\autoref{supp-rarefaction}). Sequence clustering at a similarity level
of 97 \si{\percent} yielded a total of 8,360 different OTUs. Taxonomic
classification of reads allowed for a macrophyte associated epiphytic
community determination that was mainly composed of:
\emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
\emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.4 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

@@

Processing and visualization of 16S rRNA, metagenomic and metaproteomic
data was done using R (version 3.6.0) (R Core Team, 2019), package
tidyverse (version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple
other packages (Xie, 2014, 2015, 2019, 2020; Neuwirth, 2014; Xie
\emph{et al.}, 2018; Wilke, 2018; Allaire \emph{et al.}, 2019; Zhu,
2019; Bengtsson, 2020). The detailed analysis procedure including the R
Markdown file for this paper are available as a GitHub repository
(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
Surfaces}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

@@

@@

To evaluate if the obtained DNA is suitable to determine the microbial
community structure an Illumina sequencing of the V4 16S rRNA region was
performed. Sequencing yielded a total of 336,944 sequences after quality
curation and exclusion of eukaryotic, mitochondrial and no relative
sequences. The number of sequences classified as chloroplast was 97,334.
After excluding these sequences the total number of retrieved reads
dropped to 239,610, ranging from 22,596 to 52,930 sequences per sample
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off which is commonly
observed in high-throughput 16S rRNA amplicon sequencing procedures
(\autoref{supp-rarefaction}). Sequence clustering at a similarity level
of 97 \si{\percent} yielded a total of 8,355 different OTUs. Taxonomic
classification of reads allowed for a macrophyte associated epiphytic
community determination that was mainly composed of:
\emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
\emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.5 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

@@

Processing and visualization of 16S rRNA, metagenomic and metaproteomic
data was done using R (version 3.6.0) (R Core Team, 2019), package
tidyverse (version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple
other packages (Xie, 2014, 2015, 2019, 2020; Neuwirth, 2014; Xie
\emph{et al.}, 2018; Wilke, 2018; Allaire \emph{et al.}, 2019; Zhu,
2019; Bengtsson, 2020). The detailed analysis procedure including the R
Markdown file for this paper are available as a GitHub repository
(\url{https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).

### 2020-09-16 submission/supplementary.tex
Coauthor review #2.
Pipeline optimization.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplementary Information},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

@@

\title{\textbf{Supplementary Information}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
Surfaces}}
\author{}
\date{\vspace{-2.5em}}

@@

\begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
\caption{\label{tab:nseq_notus}Sample ID, sampling station, community type, sampling date, number of sequences and number of OTUs of each sample. The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus}}\\
\toprule
\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
\midrule
\endfirsthead
\caption[]{Sample ID, sampling station, community type, sampling date, number of sequences and number of OTUs of each sample. The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus} \textit{(continued)}}\\
\toprule
\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
40 & Saline & \textit{Cymodocea nodosa} (Monospecific) & 4 December 2017 & 19,176 & 942\\
41 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & 24,250 & 1,123\\
42 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & 28,387 & 3,131\\
43 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & 13,661 & 1,809\\
61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 31,734 & 2,039\\
62 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & 36,488 & 3,211\\
63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 41,871 & 3,851\\*
\end{longtable}
\endgroup{}

@@

\multicolumn{5}{l}{\textbf{Funtana, \textit{Cymodocea nodosa} (Monospecific), 14 December 2017 (sample ID: 45)}}\\

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Information},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{Supplementary Information}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
Surfaces}}
  \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

@@

@@

\begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
\caption{\label{tab:nseq_notus}Sample ID, sampling station, community type, sampling date, number of sequences and number of OTUs of each sample. The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus}}\\
\toprule
\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
\midrule
\endfirsthead
\caption[]{Sample ID, sampling station, community type, sampling date, number of sequences and number of OTUs of each sample. The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus} \textit{(continued)}}\\
\toprule
\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
40 & Saline & \textit{Cymodocea nodosa} (Monospecific) & 4 December 2017 & 19,176 & 935\\
41 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & 24,252 & 1,129\\
42 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & 28,385 & 3,125\\
43 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & 13,667 & 1,816\\
61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 31,736 & 2,050\\
62 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & 36,499 & 3,216\\
63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 41,842 & 3,848\\*
\end{longtable}
\endgroup{}

@@

\multicolumn{5}{l}{\textbf{Funtana, \textit{Cymodocea nodosa} (Mixed), 14 December 2017 (sample ID: 45)}}\\
